Combining formal and distributional models of predicate semantics

### 2016-03-11 papers/acl2016/paper.tex
Shorten some references to K&M, add a couple more results
Former-commit-id: bfe39fdd2351aa7c7d3f5eea8e44eca29266fb29

  \vspace{.1in}
  \begin{minipage}{0.5\linewidth}
    \textbf{Category models}\\
    \begin{tabular}{@{}lll}
      \predicate{architect}: & $\theta$: &[.2, -.6, \ldots] \\
      & $\omega$: & \textsc{type:architect} $\rightarrow$ .52 \\
      &           & \textsc{type:designer} $\rightarrow$ .32 \\
      &           & \textsc{nationality:Italy} $\rightarrow$ .20 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \textbf{Relation models}\\
    \begin{tabular}{@{}lll}
      \predicate{architect\_N/N}: & $\theta$: &[-.9, .1, \ldots] \\
      & $\omega$: & \textsc{/person/nationality\textsuperscript{-1}} $\rightarrow$ .29 \\
      &           & \textsc{/structure/architect} $\rightarrow$ .11 \\
      &           & \textsc{/person/ethnicity\textsuperscript{-1}} $\rightarrow$ .05 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}

@@

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} introduced a
technique for learning semantic parsers with open predicate vocabularies.
Instead of mapping text to Freebase queries, they applied the matrix
factorization semantics used by Riedel et
al.~\shortcite{riedel-2013-mf-universal-schema} in knowledge base completion to
the task of learning predicate semantics in a semantic parser.  This method
assigns truth values to instantiated predicates like
\predicate{architect}(\entity{AndreaPalladio}) in a probabilistic database.
The models Krishnamurthy and Mitchell use for the predicates and entities are
essentially equivalent to modern techniques for constructing word embeddings,
implicitly factorizing a cooccurrence matrix between categories and entities,
and relations and entity pairs~\cite{levy-2014-w2v-as-mf}.  This implicit
factorization has two sets of parameters: each predicate (category or relation)
has a learned $k$-dimensional embedding $\theta$, and each entity and entity
pair has a learned embedding $\phi$.  The probability assigned to a category
instance $c(e)$ and relation instance $r(e_1, e_2)$ in the probabilistic
database is given by:

@@

In this paper, we use the system produced by Krishnamurthy and Mitchell,
changing only the models of predicate semantics, as described in
\secref{method}.

@@

We also used the test set created by Krishnamurthy and Mitchell, which contains
220 queries generated in the same fashion as the training data from a separate
section of ClueWeb.  However, we used this set for intermediate evaluations
while developing our models, and so we also generated another, similar test set
from a different held out section of ClueWeb.  This final test set contains 307
queries.  We report results on both of these test sets below.  All of the data
used in these experiments is available at [url withheld for review].

@@

\begin{table}
  \centering
  {\small
    \begin{tabular}{lccc}
      \toprule
      Method & MAP & Weighted MAP & MRR \\
      \midrule
      Distributional-only & .284 & .371 & .379 \\
      \ \ \ (old lfs) & ?? & ?? & ?? \\
      \midrule
      Formal-only & .276 & .469 & .334 \\
      \ \ \ (old lfs) & ?? & ?? & ?? \\
      \midrule
      Combined model & \textbf{.335} & \textbf{.477} & \textbf{.429} \\
      \ \ \ (old lfs) & ?? & ?? & ?? \\
      \bottomrule
    \end{tabular}
  }
  \caption{Results on the development set for our fill-in-the-blank task.  The
  combined model improves MAP by 42\% compared to the distributional-only
  baseline.  Additionally, the improved logical forms presented in
  \secref{better-lfs} significantly improve all models.}
  \label{tab:dev-results}
\end{table}

\tabref{dev-results} shows the results of our experiments on the development
set (the same as the test set used by Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}).  As can be
seen, the combined model significantly improves performance, giving a 42\%
relative increase in MAP over the distributional-only baseline.  This table
also shows the benefit of using our modified logical forms (see
\secref{better-lfs}); by being able to better model the relations between
entities, MAP of all of the models is significantly improved.\footnote{Because
of the large amount of annotation effort involved, we did not include the
unique results from this run in the evaluation pool.  It is possible that this
had the effect of lowering the MAP score for these methods.  However, the large
majority of the predictions made by these methods were already annotated, and
spot-checking the unannotated ones showed that practically all of the unique
predictions made in this scenario were incorrect, and thus annotating them
would not change the score.}

>>>>

  \vspace{.2in}
  \begin{minipage}{0.5\linewidth}
    \textbf{Category models}\\
    \begin{tabular}{@{}lll}
      \predicate{architect}: & $\theta$: &[.2, -.6, \ldots] \\
      & $\omega$: & \textsc{type:architect} $\rightarrow$ .52 \\
      &           & \textsc{type:designer} $\rightarrow$ .32 \\
      &           & \textsc{nationality:Italy} $\rightarrow$ .20 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \textbf{Relation models}\\
    \begin{tabular}{@{}lll}
      \predicate{architect\_N/N}: & $\theta$: &[-.9, .1, \ldots] \\
      & $\omega$: & \textsc{/person/nationality\textsuperscript{-1}} $\rightarrow$ .29 \\
      &           & \textsc{/structure/architect} $\rightarrow$ .11 \\
      &           & \textsc{/person/ethnicity\textsuperscript{-1}} $\rightarrow$ .05 \\
      &           & $\cdots$
    \end{tabular}
  \end{minipage}

@@

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} introduced a
technique for learning semantic parsers with open predicate vocabularies
(hereafter referred to as K\&M).  Instead of mapping text to Freebase queries,
they applied the matrix factorization semantics used by Riedel et
al.~\shortcite{riedel-2013-mf-universal-schema} in knowledge base completion to
the task of learning predicate semantics in a semantic parser.  This method
assigns truth values to instantiated predicates like
\predicate{architect}(\entity{AndreaPalladio}) in a probabilistic database.
The models K\&M use for the predicates and entities are essentially equivalent
to modern techniques for constructing word embeddings, implicitly factorizing a
cooccurrence matrix between categories and entities, and relations and entity
pairs~\cite{levy-2014-w2v-as-mf}.  This implicit factorization has two sets of
parameters: each predicate (category or relation) has a learned $k$-dimensional
embedding $\theta$, and each entity and entity pair has a learned embedding
$\phi$.  The probability assigned to a category instance $c(e)$ and relation
instance $r(e_1, e_2)$ in the probabilistic database is given by:

@@

In this paper, we use the system produced by K\&M, changing only the models of
predicate semantics, as described in \secref{method}.

@@

We also used the test set created by K\&M, which contains
220 queries generated in the same fashion as the training data from a separate
section of ClueWeb.  However, we used this set for intermediate evaluations
while developing our models, and so we also generated another, similar test set
from a different held out section of ClueWeb.  This final test set contains 307
queries.  We report results on both of these test sets below.  All of the data
used in these experiments is available at [url withheld for review].

@@

\begin{table}
  \centering
  {\small
    \begin{tabular}{lccc}
      \toprule
      Method & MAP & Weighted MAP & MRR \\
      \midrule
      Distributional-only & .284 & .371 & .379 \\
      \ \ \ (old lfs) & .269 & .330 & .380 \\
      \midrule
      Formal-only & .276 & .469 & .334 \\
      \ \ \ (old lfs) & .231 & .358 & .216 \\
      \midrule
      Combined model & \textbf{.335} & \textbf{.477} & \textbf{.429} \\
      \ \ \ (old lfs) & .313 & .396 & .413 \\
      \bottomrule
    \end{tabular}
  }
  \caption{Results on the development set for our fill-in-the-blank task.  The
  combined model improves MAP by 42\% compared to the distributional-only
  baseline.  Additionally, the improved logical forms presented in
  \secref{better-lfs} significantly improve all models.}
  \label{tab:dev-results}
\end{table}

\tabref{dev-results} shows the results of our experiments on the development
set (the same as the test set used by Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}).  As can be
seen, the combined model significantly improves performance, giving a 42\%
relative increase in MAP over the distributional-only baseline.  This table
also shows the benefit of using our modified logical forms (see
\secref{better-lfs}); by being able to better model the relations between
entities, MAP of all of the models is improved.  Not surprisingly, the
improvement is more pronounced for the models that have formal features, which
can make better use of rich information from Freebase with our improved logical
forms.\footnote{Because of the large amount of annotation effort involved, we
did not include the unique results from this run in the evaluation pool.  This
has the effect of slightly lowering the MAP score for these methods.  However,
the large majority of the predictions made by these methods were already
annotated, and spot-checking the unannotated predictions seemed to indicate
that annotating them would not substantially change the result shown.}
